Young children rationally integrate semantic and pragmatic information sources to infer the meaning of words

### 2020-09-27 paper/manuscript.Rmd
first draft paper

Humans rely on language for communication, making language acquisition a key learning objective of early childhood [@tomasello2018natural]. A central problem in language learning is referent identification: To learn the conventional symbolic relation between a word and an object, the child has to determine the intended referent of the word. Referents can only be identified in an inferential manner by reasoning about the speaker’s intentions [@bohn2019pervasive; @bruner1983child; @clark2009first; @tomasello2009constructing]. That is, the child has to infer what the speaker is communicating about based on information sources in the social context of the utterance.

@@

We focus on three information sources that play an integral part in theorizing about language use and learning: (1) expectations that speakers communicate in a cooperative and informative manner [@frank2014inferring; @clark1987principle], (2) shared common ground about what is being talked about in conversation [@bohn2018common; @clark1996using; @clark2015common], and (3) semantic knowledge about previously learned word–object mappings [@fenson1994variability; @bloom2002children]. These information sources operate on different timescales: while (1) is a momentary expectation about a particular utterance, (2) grows over the course of a conversation, and (3) is learned across development. This interplay of timescales has been hypothesized to be an important component of word meaning inference [@frank2009using; @mcmurray2012word].

@@

To test the predictive and explanatory power of our model, we developed a world-learning paradigm in which we jointly manipulated the three information sources (Fig. 1). Children interacted on a tablet computer with a series of storybook speakers [@frank2016using]. The speakers used novel words (e.g., "wug") in the context of two potential referents. This situation is depicted in Fig. 1a iv, in which a speaker (here, a frog) appears with a known object (a duck, left) and an unfamiliar object (the diamond-shaped object, right). To identify the referent, the child could rely upon the three information sources described above. First, they may infer that a cooperative speaker would have used the word "duck" to refer to the known object (the duck); the fact that the speaker did not say duck then suggests that the speaker is most likely referring to a different object (the unfamiliar object). This basic inference is oftentimes referred to as a mutual exclusivity inference [@markman1988children; @halberda2003development]. More generally speaking, the listener can infer the referent of the novel word by assuming that the speaker is a rational agent who chooses the most informative words to refer to objects [@clark1987principle; @frank2014inferring]. Second, the child may rely upon what has already been established in the common ground with the speaker. Listeners expect speakers to communicate about things that are new to the common ground [@akhtar1996role; @diesendruck2004two]. Thus, the inference about the novel word referring to the unfamiliar object also depends on which object is new in context (Fig. 1a and b i-iii). Finally, the child may rely upon their previously acquired semantic knowledge, that is, how sure they are that the known object is called "duck". If the known object is something less familiar, such as a chess piece (e.g., a pawn), a 3-year-old child may draw a weaker inference, if they draw any inference at all [@lewis2020-me]. Taken together, the child has the opportunity to integrate their assumptions about (1) informative communication, (2) their understanding of the common ground, and (3) their existing semantic knowledge. In some versions of the experiment, some of these information sources were aligned (Fig. 1a) but in some they were in conflict (Fig. 1b).

@@

We tested the model in its ability to predict children’s actual behavior - at an age (between 2 and 5 years) at which they learn language at an exponential rate [@bloom2002children]. We estimated children’s (N=148) developing sensitivity to individual information sources in separate experiments (Experiment 1 and 2) (see Fig. 1c-e). We then generated parameter-free, quantitative, *a priori* model predictions (developmental trajectories) representing the model's expectations about how children should behave in a new situation in which all three information sources had to be integrated. We generated model predictions for 24 experimental conditions: 12 objects of different familiarities (requiring different levels of semantic knowledge), with novelty either conflicting or coinciding; Fig. 1. We compared these predictions to newly collected data from N = 220 children from the same age range (Experiment 3). The results showed a very close alignment between model predictions and the data across the entire age range. That is, the average developmental trajectories predicted by the model resembled the trajectories found in the data (Fig. S6). For a more quantitative analysis, we binned predictions and data by year and correlated them. We found a high correlation, with the model explaining 79% of the variance in the data (Fig. 2a). These results seem to support the assumptions in the model that children integrate all available information sources. However, it is still possible that simpler models might make equally good -- or even better -- predictions.

@@

The fact that children integrated all the available information sources may be surprising, but does not rule out that the integration process itself might be different from the one described in the rational integration model. Remember that this model assumes that all information sources enter into an integrated inference process. However, a more modular view seems equally plausible: children might compute separate inferences based on a subset of the available information and then integrate them by weighting them according to some ratio. This view would allow for the possibility that some information sources are considered to be more important than others. In other words, children might be biased towards some information sources. We formalized this alternative view as a *biased integration model*. This model assumes that semantic knowledge and expectations about speaker informativeness enter into one inference (mutual exclusivity inference [@markman1988children; @lewis2020-me; @clark1987principle]) while common ground information enters into a second one. The outcome of both processes are then weighted according to the parameter $\phi$. Like the rational integration model, this model takes in all available information sources in an age sensitive way and assumes that they are integrated - the integration process itself, however, is construed in a different way.

The parameter $\phi$ in the biased integration model is unknown ahead of time and has to be estimated based on the experimental data. That is, in contrast to the rational integration model, the biased integration model does not allow to make *a priori* predictions about new data in the way we describe above. For the model comparison, we therefore constrained the parameters in the rational integration model by the experimental data as well. As a consequence, both models use all the information available to constrain their parameters (see Fig. S4).

@@

The biased integration model makes reasonable predictions that explain a large proportion of the variance in the data (Fig. 3b). The parameter $\phi$ - indicating the bias to one of the inferences - was estimated to favor the mutual exclusivity inference (mode = `r phi%>%pull(mode)`; 95% highest density interval (HDI): `r phi%>%pull(lci)` - `r phi%>%pull(uci)`, see Fig. 3d). However, the rational integration model presented a much better fit to the data, both in terms of correlation and marginal likelihood of the data (Fig. 3). Thus, a fully integrated - as opposed to a segmented and biased - integration process explained the data better.

@@

The developmental bias model explains a substantial portion of the variance in the data (Fig. 3c). The estimated developmental trajectory for the bias parameter $\phi$ suggests that younger children put a higher emphasis on common ground information, while older children rely more on the mutual exclusivity inference (Fig. 3d). The relative importance of the two inferences seems to switch at around age 3. Yet again, when we directly compare the competitor models, we find that the data is several orders of magnitude more likely under the rational integration model (Fig. 3). Interestingly, when we compared the rational integration model with its parameters not constrained by the experimental data (i.e. the "pure" prediction model described above), it also outperformed both bias models (see Fig. S10 and Tab. S6). Taken together, these results show that the integration process is best described as fully integrated and stable over time.

@@

Our model is based on classic social- pragmatic views on language use and comprehension [@clark1996using; @grice1991studies; @sperber2001relevance; @tomasello2009constructing]. As a consequence, instead of assuming that different information sources feed into separate word-learning mechanisms, we assume that they all play a functional role in an integrated social inference process. Our model goes beyond previous theoretical and empirical work by specifying the computations that may underlie this inference process. Furthermore, we present a substantive theory about how this integration process develops: We assume that children become increasingly sensitive to different information sources but that the way these information sources are integrated, remains the same. We used this model to predict and explain children's information integration in a new word learning paradigm in which they had to integrate (1) their assumptions about informative communication, (2) their understanding of the common ground, and (3) their existing semantic knowledge.

@@

There are multiple ways in which the model and the experiments presented here could be extended. First, it would be important to replicate our findings in a more diverse sample, that is, with children from a different cultural background who learn a different language. In such a study, we would not expect that our results replicate in a strict sense; that is, we would not expect to see the same developmental trajectories in all cultures and/or languages. Substantial variation is much more likely. Our computational model provides a way to think about how to reconcile cross-cultural variation with a universal cognitive architecture: We expect differences in how sensitive children are to the individual information sources at different ages, but similarities in how information is integrated [@bohn2019pervasive]. In computational terms, we assume a universal architecture that specifies the relation between a set of varying parameters.

Second, it would be useful to flesh out the cognitive processes that underlie reasoning about common ground. In its current form, our model does not specify how the listener arrives at the expectation that some objects are more likely referents because they are new in common ground. We assume that common ground changes speaker's expectations about what are likely referents [@hanna2003effects] and measured these expectations to inform the parameter values in our model. Computationally, the model does not differentiate between common ground information and other reasons that might make an object contextually more salient. A good starting point to address this issue could be work that formalizes reasoning about common ground as trading-off between (differing) perspectives [@heller2016perspective; @mozuraitis2018modeling].

@@

To analyze how the manipulations in each experiment affected children's behavior, we used generalized linear mixed models. Since the focus of the paper is on how information sources were integrated, we discuss these models in the supplementary material and focus here on the cognitive models instead. A detailed, mathematical description of the different cognitive models along with details about estimation procedures and priors can be found in the supplementary material. All Bayesian cognitive models were implemented in the probabilistic programming language `WebPPL` [@dippl]. The corresponding model code can be found in the associated online repository.

>>>>

Human communicative communicative abilities are unrivaled in the animal kingdom [@taylor2016language; @tomasello2008origins; @fitch2010evolution]. Language - in whatever modality - is the medium that allows humans to collaborate and coordinate in species-unique ways, making it the bedrock of human culture and society [@smith1997major]. Thus, in order to absorb the culture around them and become functioning members of a society, children need to learn language [@tomasello2018natural]. A central problem in language learning is referent identification: to learn the conventional symbolic relation between a word and an object, a child has to determine the intended referent of the word. Referents can only be identified in an inferential manner by reasoning about the speaker’s intentions [@bohn2019pervasive; @bruner1983child; @clark2009first; @tomasello2009constructing]. That is, the child has to infer what the speaker is communicating about based on information sources in the social context of the utterance.

@@

We focus on three information sources that play an integral part in theorizing about language use and learning: (1) expectations that speakers communicate in a cooperative and informative manner [@frank2014inferring; @clark1987principle; @sperber2001relevance], (2) shared common ground about what is being talked about in conversation [@bohn2018common; @clark1996using; @clark2015common], and (3) semantic knowledge about previously learned word–object mappings [@fenson1994variability; @bloom2002children]. These information sources operate on different timescales: while (1) is a momentary expectation about a particular utterance, (2) grows over the course of a conversation, and (3) is learned across development. This interplay of timescales has been hypothesized to be an important component of word meaning inference [@frank2009using; @mcmurray2012word].

@@

To test the predictive and explanatory power of our model, we developed a world-learning paradigm in which we jointly manipulated the three information sources (Fig. 1). Children interacted on a tablet computer with a series of storybook speakers [@frank2016using]. The speakers used novel words (e.g., "wug") in the context of two potential referents. This situation is depicted in Fig. 1a iv, in which a speaker (here, a frog) appears with a known object (a duck, left) and an unfamiliar object (the diamond-shaped object, right). To identify the referent, the child could rely upon the three information sources described above. First, they may infer that a cooperative speaker would have used the word "duck" to refer to the known object (the duck); the fact that the speaker did not say duck then suggests that the speaker is most likely referring to a different object (the unfamiliar object). This basic inference is oftentimes referred to as a mutual exclusivity inference [@markman1988children; @halberda2003development]. More generally speaking, the listener can infer the referent of the novel word by assuming that the speaker is a rational agent who chooses the most informative words to refer to objects [@clark1987principle; @frank2014inferring]. Second, the child may rely upon what has already been established in the common ground with the speaker. Listeners expect speakers to communicate about things that are new to the common ground [@akhtar1996role; @diesendruck2004two]. Thus, the inference about the novel word referring to the unfamiliar object also depends on which object is new in context (Fig. 1a and b i-iii). Finally, the child may rely upon their previously acquired semantic knowledge, that is, how sure they are that the known object is called "duck". If the known object is something less familiar, such as a chess piece (e.g., a pawn), a 3-year-old child may draw a weaker inference, if they draw any inference at all [@lewis2020-me; @grassmann2015children]. Taken together, the child has the opportunity to integrate their assumptions about (1) cooperative communication, (2) their understanding of the common ground, and (3) their existing semantic knowledge. In some versions of the experiment, some of these information sources were aligned (Fig. 1a) but in some they were in conflict (Fig. 1b).

@@

We tested the model in its ability to predict children’s actual behavior - at an age (between 2 and 5 years) at which they learn language at an exponential rate [@bloom2002children]. We estimated children’s (N=148) developing sensitivity to individual information sources in separate experiments (Experiment 1 and 2) (see Fig. 1c-e). We then generated parameter-free, quantitative, *a priori* model predictions (developmental trajectories) representing the model's expectations about how children should behave in a new situation in which all three information sources had to be integrated. We generated model predictions for 24 experimental conditions: 12 objects of different familiarities (requiring different levels of semantic knowledge), with novelty either conflicting or coinciding; Fig. 1. We compared these predictions to newly collected data from N = 220 children from the same age range (Experiment 3).

The results showed a very close alignment between model predictions and the data across the entire age range. That is, the average developmental trajectories predicted by the model resembled the trajectories found in the data (Fig. S6). For a more quantitative analysis, we binned predictions and data by year and correlated them. We found a high correlation, with the model explaining 79% of the variance in the data (Fig. 2a). These results seem to support the assumptions in the model that children integrate all available information sources. However, it is still possible that simpler models might make equally good -- or even better -- predictions.

@@

The fact that children integrated all the available information sources may be surprising, but does not rule out that the integration process itself might be different from the one described in the rational integration model, which assumes that all information sources enter into an integrated inference process. However, a more modular view seems equally plausible: children might compute separate inferences based on a subset of the available information and then integrate them by weighting them according to some ratio. This view would allow for the possibility that some information sources are considered to be more important than others. In other words, children might be biased towards some information sources. We formalized this alternative view as a *biased integration model*. This model assumes that semantic knowledge and expectations about speaker informativeness enter into one inference (mutual exclusivity inference [@markman1988children; @lewis2020-me; @clark1987principle]) while common ground information enters into a second one. The outcome of both processes are then weighted according to the parameter $\phi$. Like the rational integration model, this model takes in all available information sources in an age sensitive way and assumes that they are integrated - the only difference lies in the nature of the integration process.

The parameter $\phi$ in the biased integration model is unknown ahead of time and has to be estimated based on the experimental data. That is, in contrast to the rational integration model, the biased integration model does not allow to make *a priori* predictions about new data in the way we describe above. For the model comparison, we therefore constrained the parameters in the rational integration model by the data from EXperiment 3 as well. As a consequence, both models use all the information available to constrain their parameters (see Fig. S4).

@@

The biased integration model makes reasonable predictions that explain a large proportion of the variance in the data (Fig. 3b). The parameter $\phi$ - indicating the bias to one of the inferences - was estimated to favor the mutual exclusivity inference (mode = `r phi%>%pull(mode)`; 95% highest density interval (HDI): `r phi%>%pull(lci)` - `r phi%>%pull(uci)`, see Fig. 3d). However, the rational integration model presented a much better fit to the data, both in terms of correlation and marginal likelihood of the data (Fig. 3). Thus, a fully integrated - as opposed to a sequential and biased - integration process explained the data better.

@@

The developmental bias model also explains a substantial portion of the variance in the data (Fig. 3c). The estimated developmental trajectory for the bias parameter $\phi$ suggests that younger children put a higher emphasis on common ground information, while older children rely more on the mutual exclusivity inference (Fig. 3d). The relative importance of the two inferences seems to switch at around age 3. Yet again, when we directly compare the competitor models, we find that the data is several orders of magnitude more likely under the rational integration model (Fig. 3). Interestingly, when we compared the rational integration model with its parameters not constrained by the experimental data (i.e. the "pure" prediction model described above), it also outperformed both bias models (see Fig. S10 and Tab. S6). Taken together, these results show that the integration process is best described as fully integrated and stable over time.

@@

Our work extends the study of information integration during language comprehension [@tanenhaus1995integration; @kamide2003integration; @hagoort2004integration; @ozyurek2007line; @heller2016perspective] to pragmatics. Our model is based on classic social-pragmatic theories on language use and comprehension [@clark1996using; @grice1991studies; @sperber2001relevance; @tomasello2009constructing]. As a consequence, instead of assuming that different information sources feed into separate word-learning mechanisms, we assume that they all play a functional role in an integrated social inference process. Our model goes beyond previous theoretical and empirical work by specifying the computations that may underlie this inference process. Furthermore, we present a substantive theory about how this integration process develops: We assume that children become increasingly sensitive to different information sources but that the way these information sources are integrated, remains the same. We used this model to predict and explain children's information integration in a new word learning paradigm in which they had to integrate (1) their assumptions about informative communication, (2) their understanding of the common ground, and (3) their existing semantic knowledge.

@@

There are multiple ways in which the model and the experiments presented here could be extended. First, it would be important to replicate our findings in a more diverse sample, that is, with children from a different cultural background who learn a different language [@nielsen2017persistent]. In such a study, we would not expect that our results replicate in a strict sense; that is, we would not expect to see the same developmental trajectories in all cultures and/or languages. Substantial variation is much more likely. Nevertheless, our computational model provides a way to think about how to reconcile cross-cultural variation with a universal cognitive architecture: We expect differences in how sensitive children are to the individual information sources at different ages, but similarities in how information is integrated [@bohn2019pervasive]. In computational terms, we assume a universal architecture that specifies the relation between a set of varying parameters.

Second, it would be useful to flesh out the cognitive processes that underlie reasoning about common ground. In its current form, our model does not specify how the listener arrives at the expectation that some objects are more likely referents because they are new in common ground. We assume that common ground changes speaker's expectations about what are likely referents [@hanna2003effects] and measured these expectations to inform the parameter values in our model. Computationally, the model does not differentiate between common ground information and other reasons that might make an object contextually more salient.

@@

To analyze how the manipulations in each experiment affected children's behavior, we used generalized linear mixed models. Since the focus of the paper is on how information sources were integrated, we discuss these models in the supplementary material and focus here on the cognitive models instead. A detailed, mathematical description of the different cognitive models along with details about estimation procedures and priors can be found in the supplementary material. All Bayesian cognitive models and Bayesian data analytic models were implemented in the probabilistic programming language `WebPPL` [@dippl]. The corresponding model code can be found in the associated online repository.
